### 2017-09-01 writeup/writeup.tex
compile typos fixed hopefully

\begin{equation}
\phi(z,\chi) = \frac{\int_0^\tau (z+\tilde{r}\Phi(\tau^\prime)) e^{-\theta^\prime^2/2} d\tau^\prime}{e^{-\theta^2/2}}.
\end{equation}

The $z$ term in the integrand simply becomes $z\sqrt{2\pi} \sim z$.
The second term needs to be treated with slightly more care.
The major contribution to the integral will come from a well defined, large $\tau^\prime$, corresponding to the maximum of $\exp(-\theta^\prime^2/2 + \log \Phi(\tau^\prime))$ at $\tau^\prime = \tau - \theta$.
$\Phi(\tau)$ changes slowly and can be pulled out of the integral, yielding another Gaussian integral.
We therefore have $\phi(z,\chi) \approx (z+\tilde{r}\Phi(z))e^{\theta^2}/2$, or, summarizing:

@@

\begin{equation}
\partial_t G(\lambda,t) = (1+s+x-\bar{x})\sum_n (n-1) \lambda^n p_t(n-1) + \sum_n (n+1) \lambda^n p_t(n+1) - (2+s+x-\bar{x})\sum_n n \lambda^n p_t(n).

\end{equation}

@@

\begin{equation}
\partial_t G(\lambda,t) = \left( (1+s+x-\bar{x})\lambda^2 - (2+s+x-\bar{x})\lambda + 1)\partial_\lambda G(\lambda, t).
\end{equation}

@@

The normal to this integral surface is given by
\begin{equation}
\vec{n} = \vec{\nabla} F = G_\lambda \vec{i} + G_t \vec{j} - \vec{k}.
\end{equation}
So the PDE is simply $(a(\lambda,t),b(\lambda,t),f(\lambda,t,G)) \cdot \vec{n} = 0$, which is another way of saying that the PDE's vector field $\vec{v} = a(\lambda,t)\vec{i} + b(\lambda,t)\vec{j} + f(\lambda,t,G)\vec{k}$ is normal to the integral surface at every point.
This means $\vec{v}$ can be written as a curve along the integral surface $F$.
We can parametrize it using the dummy variable $\tau$ and write
\begin{equation}
\vec{p}(\tau ) = \lambda(\tau) \vec{i} + t(\tau) \vec{j} + G(\tau ) \vec{k},
\end{equation}
or
\begin{equation}
\vec{p}^\prime = \frac{d\vec{p}}{d\tau} = \frac{d\lambda}{d\tau} \vec{i} + \frac{dt}{d\tau} \vec{j} + \frac{dG}{d\tau} \vec{k},
\end{equation}
which is normal to $\vec{n}$ and hence proportional to $\vec{v}$.
If the proportionality constant is $\kappa$, we have $\frac{d\lambda}{d\tau} = \kappa a(\lambda,t)$, $\frac{dt}{d\tau} = \kappa b(\lambda,t)$, and $\frac{dG}{d\tau} = \kappa f(\lambda,t,G)$, which gives us simply
\begin{equation}
\frac{d\lambda}{a} = \frac{dt}{b} = \frac{dG}{f}.
\end{equation}
In this case, we have $a(\lambda, t) = (1+s+x-\bar{x})\lambda^2 - (2+s+x-\bar{x})\lambda + 1)$, $b(\lambda, t) = -1$, and $c(\lambda, t) = f(\lambda,t,G) = 0$.
To avoid a divide by zero error, we can simply flip the above equation.
We end up with
\begin{equation}
\frac{d\lambda}{dt} = -((1+s+x-\bar{x})\lambda^2 - (2+s+x-\bar{x})\lambda + 1) = P(s,x,\lambda, t), \frac{dG}{dt} = 0,
\end{equation}
with the so called ``Cauchy data" (initial/boundary conditions) given above.
It remains to solve these ODEs.
It is worth noting that $P$ can be factored into
\begin{equation}
P(s,x,\lambda, t) = ((1+s+x-\bar{x})\lambda - 1)(\lambda - 1).
\end{equation}
The first equation we need to solve becomes:
\begin{equation}
\int \frac{d\lambda}{P(s,x,\lambda, t)} = \int dt
\end{equation}
Decomposing $P$ and using partial fractions allows us to write
\begin{equation}
\int \frac{d\lambda}{P(s,x,\lambda, t)} = \int d\lambda\left(\frac{A}{\lambda - \frac{1}{\alpha}} + \frac{B}{\lambda - 1}),
\end{equation}
with $\alpha = s+x-\bar{x}$, $A = \frac{1}{\frac{1}{\alpha} - 1}$, $B = -A = \frac{-1}{\frac{1}{\alpha} - 1}$.
Integrating the first term on the left yields $\frac{1}{\frac{1}{\alpha}-1}\log(\lambda - \frac{1}{\alpha})$, and the second term yields $-\frac{1}{\frac{1}{\alpha}-1}\log(\lambda - 1)$.
We thus have $\frac{1}{\frac{1}{\alpha}-1}\log(\frac{\lambda-\frac{1}{\alpha}}{\lambda-1}) + C_1$.
The right hand side becomes of course simply $t$.
Next, the equation $\frac{dG}{dt} = 0$ yields simply that $G = C_2$, or that $G$ depends on $t$ only through its dependence on $\lambda$.
At this point, if $\alpha$ were simply a constant (in the case of no fitness variance, there would be no need to consider $x$ at all, and $\alpha$ would equal $s$), we could simply proceed to choose a clever value for $C_1$ and $C_2$ and solve for $\lambda(t)$.
The Cauchy data $G(\lambda,0) = \lambda$ and $G(1,t) = 1$ would then guide us in determining the form of $t$.
However, $\alpha$ is not a constant: it is hiding a nasty time dependence (via $\bar{x} = \sigma^2 t$).
There might be a way around this, but it is not obvious.

>>>>

\begin{equation}
\phi(z,\chi) = \frac{\int_0^\tau (z+\tilde{r}\Phi(\tau^\prime)) e^{-{\theta'}^2/2} d\tau^\prime}{e^{-\theta^2/2}}.
\end{equation}

The $z$ term in the integrand simply becomes $z\sqrt{2\pi} \sim z$.
The second term needs to be treated with slightly more care.
The major contribution to the integral will come from a well defined, large $\tau^\prime$, corresponding to the maximum of $\exp(-\theta'^2/2 + \log \Phi(\tau^\prime))$ at $\tau^\prime = \tau - \theta$.
$\Phi(\tau)$ changes slowly and can be pulled out of the integral, yielding another Gaussian integral.
We therefore have $\phi(z,\chi) \approx (z+\tilde{r}\Phi(z))e^{\theta^2}/2$, or, summarizing:

@@

\begin{equation}
\partial_t G(\lambda,t) = (1+s+x-\bar{x})\sum_n (n-1) \lambda^n p_t(n-1) + \sum_n (n+1) \lambda^n p_t(n+1) - (2+s+x-\bar{x})\sum_n n \lambda^n p_t(n).
\end{equation}

@@

\begin{equation}
\partial_t G(\lambda,t) = \left( (1+s+x-\bar{x})\lambda^2 - (2+s+x-\bar{x})\lambda + 1 \right)\partial_\lambda G(\lambda, t).
\end{equation}

@@

The normal to this integral surface is given by
\begin{equation}
\vec{n} = \vec{\nabla} F = G_\lambda \vec{i} + G_t \vec{j} - \vec{k}.
\end{equation}
So the PDE is simply $(a(\lambda,t),b(\lambda,t),f(\lambda,t,G)) \cdot \vec{n} = 0$, which is another way of saying that the PDE's vector field $\vec{v} = a(\lambda,t)\vec{i} + b(\lambda,t)\vec{j} + f(\lambda,t,G)\vec{k}$ is normal to the integral surface at every point.
This means $\vec{v}$ can be written as a curve along the integral surface $F$.
We can parametrize it using the dummy variable $\tau$ and write
\begin{equation}
\vec{p}(\tau ) = \lambda(\tau) \vec{i} + t(\tau) \vec{j} + G(\tau ) \vec{k},
\end{equation}
or
\begin{equation}
\vec{p}^\prime = \frac{d\vec{p}}{d\tau} = \frac{d\lambda}{d\tau} \vec{i} + \frac{dt}{d\tau} \vec{j} + \frac{dG}{d\tau} \vec{k},
\end{equation}
which is normal to $\vec{n}$ and hence proportional to $\vec{v}$.
If the proportionality constant is $\kappa$, we have $\frac{d\lambda}{d\tau} = \kappa a(\lambda,t)$, $\frac{dt}{d\tau} = \kappa b(\lambda,t)$, and $\frac{dG}{d\tau} = \kappa f(\lambda,t,G)$, which gives us simply
\begin{equation}
\frac{d\lambda}{a} = \frac{dt}{b} = \frac{dG}{f}.
\end{equation}
In this case, we have $a(\lambda, t) = (1+s+x-\bar{x})\lambda^2 - (2+s+x-\bar{x})\lambda + 1)$, $b(\lambda, t) = -1$, and $c(\lambda, t) = f(\lambda,t,G) = 0$.
To avoid a divide by zero error, we can simply flip the above equation.
We end up with
\begin{equation}
\frac{d\lambda}{dt} = -((1+s+x-\bar{x})\lambda^2 - (2+s+x-\bar{x})\lambda + 1) = P(s,x,\lambda, t), \frac{dG}{dt} = 0,
\end{equation}
with the so called ``Cauchy data" (initial/boundary conditions) given above.
It remains to solve these ODEs.
It is worth noting that $P$ can be factored into
\begin{equation}
P(s,x,\lambda, t) = ((1+s+x-\bar{x})\lambda - 1)(\lambda - 1).
\end{equation}
The first equation we need to solve becomes:
\begin{equation}
\int \frac{d\lambda}{P(s,x,\lambda, t)} = \int dt
\end{equation}
Decomposing $P$ and using partial fractions allows us to write
\begin{equation}
\int \frac{d\lambda}{P(s,x,\lambda, t)} = \int d\lambda\left(\frac{A}{\lambda - \frac{1}{\alpha}} + \frac{B}{\lambda - 1} \right),
\end{equation}
with $\alpha = s+x-\bar{x}$, $A = \frac{1}{\frac{1}{\alpha} - 1}$, $B = -A = \frac{-1}{\frac{1}{\alpha} - 1}$.
Integrating the first term on the left yields $\frac{1}{\frac{1}{\alpha}-1}\log(\lambda - \frac{1}{\alpha})$, and the second term yields $-\frac{1}{\frac{1}{\alpha}-1}\log(\lambda - 1)$.
We thus have $\frac{1}{\frac{1}{\alpha}-1}\log(\frac{\lambda-\frac{1}{\alpha}}{\lambda-1}) + C_1$.
The right hand side becomes of course simply $t$.
Next, the equation $\frac{dG}{dt} = 0$ yields simply that $G = C_2$, or that $G$ depends on $t$ only through its dependence on $\lambda$.
At this point, if $\alpha$ were simply a constant (in the case of no fitness variance, there would be no need to consider $x$ at all, and $\alpha$ would equal $s$), we could simply proceed to choose a clever value for $C_1$ and $C_2$ and solve for $\lambda(t)$.
The Cauchy data $G(\lambda,0) = \lambda$ and $G(1,t) = 1$ would then guide us in determining the form of $t$.
However, $\alpha$ is not a constant: it is hiding a nasty time dependence (via $\bar{x} = \sigma^2 t$).
There might be a way around this, but it is not obvious.
